Supplementary material

### 2021-12-03 analysis/paper/paper.Rmd
First draft

```{r deglaciation, echo = FALSE, fig.cap = "Deglaciation at 1000 year intervals from 18-10 ka (Hughes et al. 2016). The studyarea defined later in the text is given a red outline.", out.width = "400px", fig.align="center"}

@@

To get at the relationship between sites and contemporaneous shoreline, the analysis is dependent on a study area with good control of the trajectory of prehistoric shoreline displacement. While there is displacement data available for other areas of south-eastern Norway [e.g. @hafsten1957; @sørensen1979; @sørensen1999], considerable methodological developments in recent years means that the most well-established displacement curves are from the region stretching from Halden county in the north-east, to Arendal in the south-west. This area has newly compiled displacement curves for Halden [@romundset2021], Larvik [@sørensen2014; @sørensen2014b], Tvedestrand [@romundset2018; @romundset2018b], and Arendal [@romundset2018b].

@@

anc <- as.numeric(c(bboxsites$ymin, bboxsites$xmax))
overview_map <- ggplot() +
  geom_sf(data = bmap_reproj, fill = "grey", colour = NA) +
  geom_sf(data = isobases, aes(colour = name)) +
  geom_sf(data = st_centroid(sites_sa), fill = "red",
          size = 2.25, shape = 21,
          colour = "black") +
  ggsn::scalebar(data = sites_sa, dist = 20, dist_unit = "km",
                 transform = FALSE, st.size = 4, height = 0.02,
                 border.size = 0.1, st.dist = 0.03,
                 anchor = c(x = anc[2] - 15500, y = anc[1]) + 8000) +
  coord_sf(xlim = c(bboxsites[1], bboxsites[3]),
           ylim = c(bboxsites[2], bboxsites[4]),
           expand = FALSE) +
  scale_colour_manual(values = c("Arendal" = "black",
                                   "Larvik" = "darkgreen",
                                   "Tvedestrand" = "blue",
                                   "Halden" = 'red')) +
  theme_bw() + theme(axis.title=element_blank(),
                     axis.text.y = element_blank(),
                     axis.text.x = element_blank(),
                     rect = element_rect(),
                     axis.ticks = element_blank(),
                     panel.grid.major = element_blank(),
                     legend.position = "none")

disp_curves <- ggplot(displacement_curves, aes(x = years, col = name)) +
  geom_line(aes(y =  upperelev)) +
  geom_line(aes(y = lowerelev)) +
  ylab("Meters above present sea-level") +
  xlab("cal BCE/CE") +
  scale_colour_manual(values = c("Arendal" = "black",
                                   "Larvik" = "darkgreen",
                                   "Tvedestrand" = "blue",
                                   "Halden" = 'red')) +
  theme_bw() +
  theme(legend.title = element_blank(), legend.position = "bottom",
        legend.direction = "horizontal")

>>>>

```{r deglaciation, echo = FALSE, fig.cap = "Deglaciation at 1000 year intervals from 19-10 ka (data from Hughes et al. 2016). The studyarea defined later in the text is given a red outline.", out.width = "400px", fig.align="center"}

@@

To get at the relationship between sites and contemporaneous shoreline, the analysis is dependent on a study area with good control of the trajectory of prehistoric shoreline displacement. While there is displacement data available for other areas of south-eastern Norway [e.g. @hafsten1957; @sørensen1979; @sørensen1999], considerable methodological developments in recent years means that the most well-established displacement curves are from the region stretching from Horten county in the north-east, to Arendal in the south-west. This area has newly compiled displacement curves for Horten [@romundset2021], Larvik [@sørensen2014; @sørensen2014b], Tvedestrand [@romundset2018; @romundset2018b], and Arendal [@romundset2018b].

@@

anc <- as.numeric(c(bboxsites$ymin, bboxsites$xmax))
overview_map <- ggplot() +
  geom_sf(data = bmap_reproj, fill = "grey", colour = NA) +
  geom_sf(data = isobases, aes(colour = name)) +
  geom_sf(data = st_centroid(sites_sa), fill = "red",
          size = 2.25, shape = 21,
          colour = "black") +
  ggsn::scalebar(data = sites_sa, dist = 20, dist_unit = "km",
                 transform = FALSE, st.size = 4, height = 0.02,
                 border.size = 0.1, st.dist = 0.03,
                 anchor = c(x = anc[2] - 15500, y = anc[1]) + 8000) +
  coord_sf(xlim = c(bboxsites[1], bboxsites[3]),
           ylim = c(bboxsites[2], bboxsites[4]),
           expand = FALSE) +
  scale_colour_manual(values = c("Arendal" = "black",
                                   "Larvik" = "darkgreen",
                                   "Tvedestrand" = "blue",
                                   "Horten" = 'red')) +
  theme_bw() + theme(axis.title=element_blank(),
                     axis.text.y = element_blank(),
                     axis.text.x = element_blank(),
                     rect = element_rect(),
                     axis.ticks = element_blank(),
                     panel.grid.major = element_blank(),
                     legend.position = "none")

disp_curves <- ggplot(displacement_curves, aes(x = years, col = name)) +
  geom_line(aes(y =  upperelev)) +
  geom_line(aes(y = lowerelev)) +
  ylab("Meters above present sea-level") +
  xlab("cal BCE/CE") +
  scale_colour_manual(values = c("Arendal" = "black",
                                   "Larvik" = "darkgreen",
                                   "Tvedestrand" = "blue",
                                   "Horten" = 'red')) +
  theme_bw() +
  theme(legend.title = element_blank(), legend.position = "bottom",
        legend.direction = "horizontal")

### 2021-12-03 analysis/supplementary-materials/supplementary.Rmd
First draft

```{r,  echo = FALSE, results = 'hide'}
library(here)

@@

```{r adalvestre1,  echo = FALSE, out.width = "500px"}
knitr::include_graphics(here("analysis/figures/adalvestre1.png"))

>>>>

```{r,  echo = FALSE, results = 'hide'}
library(here)
library(tidyverse)

# Read in radiocarbon dates to be plotted alongside analysis for
# each site.
rcarbon <- read.csv(here('analysis/data/raw_data/radiocarbon.csv'))
```

@@

```{r adalvestre1,  echo = FALSE, out.width = "500px"}

c14 <- filter(rcarbon, site_name  == "Adal vestre 1") %>% 
  select(lab_code, c14_bp, error, material, context)

knitr::include_graphics(here("analysis/figures/adalvestre1.png"))
```
